rough outline

### 2022-04-27 outline.Rmd
Moved some graphs and tables from the Draft2rmd into the outline

```{r, include=FALSE}
#This will allow us to filter through our data 
library(tidyverse)
library(dplyr)
#This will help us plot figures to showcase our findings
library(ggplot2)
```

@@

Let's view the first 6 rows of the data
```{r}
head(world_pop)
```
# Exploratory Data Analysis

>>>>

```{r, include=FALSE}
#This will allow us to filter through our data 
library(tidyverse)
library(dplyr)
#This will help us plot figures to showcase our findings
library(ggplot2)
#This will help us organize and display our data as necessary 
library(knitr)
library(kableExtra)
#This expands our plot uses 
library(plotly)
```

@@

Let's view the first 6 rows of the data
This dataset covers more than we need, and so we can trim it to the years we need.
```{r}
world_pop2 <- world_pop %>% filter(Year >= 1990 & Year <= 2017)
glimpse(world_pop2)

```
After selecting our desired years, we've cut the population data set nearly in half.

To get a general idea of the data, let's make some plots to see what's happening.

```{r}
p <- ggplot(deaths_df, aes(x = indoor_deaths, y = outdoor_deaths, text = paste0(country, ", ", year) )) + geom_point()
ggplotly(p)
```

This is a mess, and so we chose two countries from each continent (a high-population and a low-population country) to graph.

```{r}
#Mean total deaths from 1990-2017 of high-population countries
deaths_highpop_countries <- deaths_df %>% 
  filter(country %in% c('United States', 'Brazil', 'Ethiopia', 'Russia', 'China', 'Australia')) %>% 
  group_by(country) %>% 
  select(total_deaths) %>% 
  summarize(average_death_high = mean(total_deaths))

#deaths_highpop_countries

#Mean total deaths from 1990-2017 of high-population countries
deaths_lowpop_countries<- deaths_df %>% 
  filter(country %in% c('Canada', 'Argentina', 'Seychelles', 'Romania', 'Israel', 'Tonga')) %>% 
  group_by(country) %>% 
  select(total_deaths) %>% 
  summarize(average_death_low = mean(total_deaths))

#death_lowpop_countries
kable(list(deaths_highpop_countries, deaths_lowpop_countries))
kable(deaths_highpop_countries) %>% 
  kable_styling(full_width = FALSE, position = 'float_left')
kable(deaths_lowpop_countries) %>% 
  kable_styling(full_width = FALSE, position = 'left')

ggplot(deaths_highpop_countries)+
  geom_col(mapping = aes(x=country, y=average_death_high))+
             xlab("Country")+
             ylab("Average deaths (per 100,000)")+
             ggtitle("Average total deaths in high-population countries")+
  coord_flip()

ggplot(deaths_lowpop_countries)+
  geom_col(mapping = aes(x=country, y=average_death_low))+
             xlab("Country")+
             ylab("Average deaths (per 100,000)")+
             ggtitle("Average total deaths in low-population countries")+
  coord_flip()

```
This shows us the deaths due to pollution, but what about the average population of those countries at that time?

```{r}
hp_countries_population <- world_pop2 %>% 
  filter(Country.Name %in% c('United States', 'Brazil', 'Ethiopia', 'Russia', 'China', 'Australia')) %>% 
  group_by(Country.Name) %>% 
  select(Count) %>% 
  summarize(average_population = mean(Count))

#hp_countries_population

lp_countries_population <- world_pop2 %>% 
  filter(Country.Name %in% c('Canada', 'Argentina', 'Seychelles', 'Romania', 'Israel', 'Tonga')) %>% 
  group_by(Country.Name) %>% 
  select(Count) %>% 
  summarize(average_population = mean(Count))

#lp_countries_population

kable(list(hp_countries_population, lp_countries_population))
kable(hp_countries_population) %>% 
  kable_styling(full_width = FALSE, position = 'float_left')
kable(lp_countries_population) %>% 
  kable_styling(full_width = FALSE, position = 'left')

```

# Exploratory Data Analysis
